Component response rate variation drives stability in large complex systems

### 2018-07-28 notebook/ms.Rmd
More discussion

Here I have shown that the stability of large systems might be contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems. *Cite @Ahmadian2015 here*?

It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $\mathbf{M}$ does not necessarily increase the probability that the system will be stable (see Supplementary Information). Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability. This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig. 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.

>>>>

Here I have shown that the stability of large systems might be contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.

It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $\mathbf{M}$ does not necessarily increase the probability that the system will be stable. Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability. This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig. 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values. The mathematics underlying this shift in eigenvalue distribution have been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but have not been interpreted as rates of response of individual componetents to system perturbation.
